Optimal Instantiation of Abstract Workflows using Logic and Circuits

### 2010-01-25 publicaciones/eswc10/paperESWC10.tex
paper con correcciones

\allowdisplaybreaks
\title{Optimal Instantiation of Abstract Workflows using Logic and Circuits}
\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
\institute{Departamento de Computaci\'on \\
           Universidad Sim\'on Bol\'{\i}var \\
           Caracas 89000, Venezuela \\
           \url{{idaniel,bonet,mvidal}@ldc.usb.ve}}
\maketitle

@@

The typical SOA consists of two layers. The concrete layers that is
made of concrete services whose functionalities are described in
terms of pre- and post-conditions and non-functional properties in
terms of QoS parameters, and the abstract layer that is made of 
software applications whose functionalities are described in terms
of abstract workflows and non-functional properties in terms of
restrictions on QoS.
The execution of an abstract workflow involves the instantiation
of the abstract services into concrete services that meet the
functional and non-functional requirements. This instantiation
process can be seen as a search for a target instantiation on
the combinatorial space of all valid instantiations.
Thus, one is interested in efficient techniques for performing
this search that are able to scale up as the number of concrete
services or the complexity of the workflow increases.
We call the problem of instantiating a given abstract workflow
with concrete services, from a given pool of concrete services,
such that certain QoS demands are meet, the Workflow Instantiation
Problem (WIP).

@@

The recent approach of Arvelo et al.\ \cite{arvelo:aaai06} is based 
on the efficient enumeration of models of a propositional logic theory. 
Given a QRP, a logical theory is constructed such that each model of
the theory encodes a valid rewriting, and thus all valid rewritings
are obtained by enumerating all models of the theory. This enumeration
can be efficiently performed if the logical theory is in certain 
normal form called deterministic and decomposable negation normal
form (d-DNNF) \cite{darwiche:d-dnnfs}. Thus, the approach consists in
transforming (called compiling in the field of knowledge compilation)
the logical theory into d-DNNF format for enumerating its models efficiently.

@@

Ko et al.\ \cite{myoung08} propose a constraint-based approach that encodes the
non-functional permissible values as a set of constraints whose violation needs
to be minimized; to traverse the space of possibly optimal solutions, a hybrid
algorithm that combines tabu search and simulated annealing meta-heuristics is
implemented. Experimental results show that the proposed solution is able to
scale up to a large number of services and abstract processes.
Cardellini et al.\ \cite{cardellini07} encode one part of the QoS-aware service
composition problem as a Linear Programming problem \cite{cardellini07}.
On the other hand, Wada el at.\ \cite{Hiroshi2008} treat the problem as a 
multi-objective optimization problem where the different QoS parameters are
considered equally important instead of aggregating them into a single function.
Then, a genetic-based algorithm is proposed to identify a set of non-dominated
service compositions that best fit all the QoS requirements.

@@

The \minicon\ algorithm overcomes the limitations of the previous algorithms by identifying
only views that rewrite a set of the sub-goals of the query, and that can be combined with
the rest of the sub-goals. The key idea is to identify the mappings between the variables in
each sub-goal to the variables in one or more sub-goals in the views, in a way
that, join variables in the query are mapped to join variables in the body of a view or to
the distinguished variables of the view. Mappings between variables and sub-goals are
represented in the so-called MiniCon Descriptions (MCDs) \cite{pottinger:minicon}.

@@

A Negation Normal Form (NNF) theory is constructed from literals using only conjunctions
and disjunctions \cite{barwise:handbook}, and it can be represented as a directed
acyclic graph in which the leaves are labeled with literals and the internal nodes
are labeled with $\land$ and $\lor$; see Fig.~\ref{fig:dnnf} for an example.
This fragment is universal meaning that for every logical formula there is an
equivalent one in NNF format.
An NNF is said to be decomposable (DNNF) \cite{darwiche:d-dnnfs} if for each conjunction
$\bigwedge_i\phi_i$, the set of variables in each conjunct are pairwise disjoint; i.e,.
$Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i\neq j$.
A DNNF supports a number of operations in polytime in the size of its DAG.
For example, we can test whether a DNNF is satisfiable by a single bottom-up pass
over its DAG in linear time.
A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each
disjunction $\bigvee_i\phi_i$, the disjuncts are pairwise logically contradictory;
i.e., $\phi_i\land\phi_j\equiv\textbf{false}$ for $i\neq j$.
The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable and deterministic.
A d-DNNF supports model counting in polytime in the size of its DAG, and model
enumeration in polytime in the size of the output.
Furthermore, given a literal ranking function $r$, one can compute the rank
of the best model in polytime for DNNFs \cite{darwiche:weighted}.

The fragments DNNF and d-DNNF are universal yet translating a CNF theory 
to DNNF format has an exponential cost in the worst case. This translation
is referred to as compilation in the field. There is a publicly available
compiler, called c2d,\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}
that perform this compilation process and that make
use of modern SAT techniques such as conflict-directed backtracking,
clause learning and caching \cite{darwiche:compiler}.
This compiler incurs in the worst case in exponential space in a parameter
called the width of the decomposition tree that is related to the ``connectivity''
of the CNF theory. However, in our experiments, the CNF theories that are
compiled are of low width.

@@

An input instance of WIP is encoded as a CNF theory whose models corresponds to
the instantiations of the workflow by the Encoder. The compiler c2d, an off-the-shelf
component, compiles the CNF formula into d-DNNF.
The Encoder translates the WIP instances into CNF theories, that are then converted
into d-DNNF using c2d. The Finder computes a best model given the QoS parameters
in linear time in the size of the resulting d-DNNF. It is important to remark,
that the compilation process needs to be performed only once as it does not
depend on the value of the QoS parameters. Thus, even if the compilation happens
to be costly in terms of time, this cost can be amortized since the resulting
d-DNNF can be used to find best instantiations with respect to multiple values
of the QoS parameters.
Finally, the Decoder translates the best model returned by the Finder into
a workflow instantiation that solves the WIP.

@@

The following clauses encode the WIP problem in terms of the WIDs.
Rajaraman et al.\ \cite{RajaramanSU95} showed that for queries without negation or
arithmetic comparisons, but with constants, and $m$ goals and $k$ variables in the
head of the workflow, it is enough to consider instantiations of length at most $N=m+k$.
\begin{enumerate}[C10.]
\item[C1.] (At least of view is used): $\bigvee_{i=0}^n v_i$.
\item[C2.] (At most one view is used): $\neg v_i\lor\neg v_j$ for $i\neq j$.
\item[C3.] (Null view equals null): $v_0 \Rightarrow \neg g_j$ for $1\leq j\leq m$.
\item[C4.] (Views are useful): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ for $1\leq i\leq n$.
\item[C5.] (Subgoals covered at most once): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ for appropriate $i,j,k,l$.
\item[C6.] (Scope of views): $v_i \Rightarrow \neg g_j$ for goals that cannot be covered by $V_i$.
\item[C7.] (Consistency): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ for appropriate $i,j,k$.
\item[C8.] (Dead variables): $v_i \Rightarrow \neg t_{x,y}$ for all $x,y$ with $y\notin V_i$.
\item[C9.] (1-1 for $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ for all
           existential variables $y,y'\in V_i$.
\item[C10.] (Distinguished): $v_i \Rightarrow \neg t_{x,y}$ for distinguished $x$ and existential
            $y\in V_i$.
\item[C11.] (Existential): $v_i\land t_{x,y}\Rightarrow g_j$ for exist.\ $y\in V_i$ and
            goals $g_j$ that contain $x$.
\item[C12.] (Match): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ for all $x,y$ that must match
            if $g_j$ is covered by goal $k$ in $V_i$.
\item[C13.] (If all vars in $V_i$ are distinguished, it covers only one goal):
            $v_i \land g_j \Rightarrow \neg g_k$ for appropriate views $v_i$.
\end{enumerate}
These clauses are the same clauses used by \mcdsat\ to encode QRPs \cite{arvelo:aaai06}.
In order to properly manage constant symbols, the clauses must be enhanced with:
\begin{enumerate}[C10.]
\item[C14.] (Direct inconsistency 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
\item[C15.] (Direct inconsistency 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
\item[C16.] (Direct inconsistency 3): $\neg t_{A,B}$.
\item[C17.] (Transitivity 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
\item[C18.] (Transitivity 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
\end{enumerate}
Recall that, as shown in Section~2, the main issue when handling
constants is to be sure that not two different constants are mapped
into each other either directly or indirectly.
Clauses C14--C16 prune direct inconsistent mappings, while the last
two clauses implement a restricted propagation of mappings that prune
indirect inconsistencies.
Likewise, the theory $\Delta_{com}$ that specifies how a complete
instantiation contains the following clauses:
\begin{enumerate}[C10.]
\item[C19.] (Cover all goals): $\bigvee_{j=1}^m g^i_j$ for $1\leq i\leq N$.
\item[C20.] (Disjunctive cover): $g^i_k \Rightarrow \neg g^j_k$ for $i\neq j$.
\item[C21.] (Prune symmetries): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
            for $1\leq j\leq m$ and $1\leq i\leq N$.
\item[C22.] (Direct inconsistency 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
\end{enumerate}
These clauses provide sound and complete characterization of WIPs in the
sense that their models are in correspondence with the
instantiations of the workflow.

>>>>

\allowdisplaybreaks
\title{Optimal Instantiation of Abstract Workflows using Logic and Circuits}
\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
\institute{Departamento de Computaci\'on \\
           Universidad Sim\'on Bol\'{\i}var \\
           Caracas 89000, Venezuela \\
           \url{{idaniel,mvidal,bonet}@ldc.usb.ve}}
\maketitle

@@

The typical SOA consists of two layers. The concrete layer that is
made of concrete services whose functionalities are described in
terms of pre- and post-conditions and non-functional properties in
terms of QoS parameters, and the abstract layer that is made of 
software applications whose functionalities are described in terms
of abstract workflows and non-functional properties in terms of
restrictions on QoS.
The execution of an abstract workflow involves the instantiation
of the abstract services into concrete services that meet the
functional and non-functional requirements. This instantiation
process can be seen as a search for a target instantiation on
the combinatorial space of all valid instantiations.
Thus, one is interested in efficient techniques for performing
this search that are able to scale up as the number of concrete
services or the complexity of the workflow increases.
We call the problem of instantiating a given abstract workflow
with concrete services, from a given pool of concrete services,
such that certain QoS demands are meet, the Workflow Instantiation
Problem (WIP).

@@

The recent approach of Arvelo et al.\ \cite{arvelo:aaai06} is based 
on the efficient enumeration of models of a propositional logic theory. 
Given a QRP, a logical theory is constructed such that each model of
the theory encodes a valid rewriting, and thus all valid rewritings
are obtained by enumerating all models of the theory. This enumeration
can be efficiently performed if the logical theory is in a certain 
normal form called deterministic and decomposable negation normal
form (d-DNNF) \cite{darwiche:d-dnnfs}. Thus, the approach consists in
transforming (called compiling in the field of knowledge compilation)
the logical theory into d-DNNF format for enumerating its models efficiently.

@@

Ko et al.\ \cite{myoung08} propose a constraint-based approach that encodes the
non-functional permissible values as a set of constraints whose violation needs
to be minimized; to traverse the space of possibly optimal solutions, a hybrid
algorithm that combines tabu search and simulated annealing meta-heuristics is
implemented. Experimental results show that the proposed solution is able to
scale up to a large number of services and abstract processes.
Cardellini et al.\ \cite{cardellini07} encode one part of the QoS-aware service
composition problem as a Linear Programming problem \cite{cardellini07}.
On the other hand, Wada et al.\ \cite{Hiroshi2008} treat the problem as a 
multi-objective optimization problem where the different QoS parameters are
considered equally important instead of aggregating them into a single function.
Then, a genetic-based algorithm is proposed to identify a set of non-dominated
service compositions that best fit all the QoS requirements.

@@

The \minicon\ algorithm overcomes the limitations of the previous algorithms by identifying
only views that rewrite a set of the sub-goals of the query, and that can be combined with
the rest of the sub-goals. The key idea is to identify the mappings between the variables in
each sub-goal to the variables in one or more sub-goals in the views, in a way
that join variables in the query are mapped to join variables in the body of a view or to
the distinguished variables of the view. Mappings between variables and sub-goals are
represented in the so-called MiniCon Descriptions (MCDs) \cite{pottinger:minicon}.

@@

A Negation Normal Form (NNF) theory is constructed from literals using only conjunctions
and disjunctions \cite{barwise:handbook}, and it can be represented as a directed
acyclic graph in which the leaves are labeled with literals and the internal nodes
are labeled with $\land$ and $\lor$; see Fig.~\ref{fig:dnnf} for an example.
This fragment is universal meaning that for every logical formula there is an
equivalent one in NNF format.
An NNF is said to be decomposable (DNNF) \cite{darwiche:d-dnnfs} if for each conjunction
$\bigwedge_i\phi_i$, its variables are pairwise disjoint; i.e,.
$Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i\neq j$.
A DNNF supports a number of operations in polytime in the size of its DAG.
For example, we can test whether a DNNF is satisfiable by a single bottom-up pass
over its DAG in linear time.
A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each
disjunction $\bigvee_i\phi_i$, the disjuncts are pairwise logically contradictory;
i.e., $\phi_i\land\phi_j\equiv\textbf{false}$ for $i\neq j$.
The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable and deterministic.
A d-DNNF supports model counting in polytime in the size of its DAG, and model
enumeration in polytime in the size of the output.
Furthermore, given a literal ranking function $r$, one can compute the rank
of the best model in polytime for DNNFs \cite{darwiche:weighted}.

The fragments DNNF and d-DNNF are universal yet translating a CNF theory 
to DNNF format has an exponential cost in the worst case. This translation
is referred to as compilation in the field. There is a publicly available
compiler, called c2d,\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}
that performs this compilation process and that makes
use of modern SAT techniques such as conflict-directed backtracking,
clause learning and caching \cite{darwiche:compiler}.
This compiler incurs in the worst case in exponential space in a parameter
called the width of the decomposition tree that is related to the ``connectivity''
of the CNF theory. However, in our experiments, the CNF theories that are
compiled are of low width.

@@

An input instance of WIP is encoded as a CNF theory whose models correspond to
the instantiations of the workflow by the Encoder. The compiler c2d, an off-the-shelf
component, compiles the CNF formula into d-DNNF.
The Encoder translates the WIP instances into CNF theories, that are then converted
into d-DNNF using c2d. The Finder computes a best model given the QoS parameters
in linear time in the size of the resulting d-DNNF. It is important to remark
that the compilation process needs to be performed only once as it does not
depend on the value of the QoS parameters. Thus, even if the compilation happens
to be costly in terms of time, this cost can be amortized since the resulting
d-DNNF can be used to find best instantiations with respect to multiple values
of the QoS parameters.
Finally, the Decoder translates the best model returned by the Finder into
a workflow instantiation that solves the WIP.

@@

The following clauses encode the WIP problem in terms of the WIDs.
Rajaraman et al.\ \cite{RajaramanSU95} showed that for queries without negation or
arithmetic comparisons, but with constants, and $m$ goals and $k$ variables in the
head of the workflow, it is enough to consider instantiations of length at most $N=m+k$.
\begin{enumerate}[C10.]
\item[C1.] (At least one view is used): $\bigvee_{i=0}^n v_i$.
\item[C2.] (At most one view is used): $\neg v_i\lor\neg v_j$ for $i\neq j$.
\item[C3.] (Null view equals null): $v_0 \Rightarrow \neg g_j$ for $1\leq j\leq m$.
\item[C4.] (Views are useful): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ for $1\leq i\leq n$.
\item[C5.] (Subgoals covered at most once): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ for appropriate $i,j,k,l$.
\item[C6.] (Scope of views): $v_i \Rightarrow \neg g_j$ for goals that cannot be covered by $V_i$.
\item[C7.] (Consistency): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ for appropriate $i,j,k$.
\item[C8.] (Dead variables): $v_i \Rightarrow \neg t_{x,y}$ for all $x,y$ with $y\notin V_i$.
\item[C9.] (1-1 for $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ for all
           existential variables $y,y'\in V_i$.
\item[C10.] (Distinguished): $v_i \Rightarrow \neg t_{x,y}$ for distinguished $x$ and existential
            $y\in V_i$.
\item[C11.] (Existential): $v_i\land t_{x,y}\Rightarrow g_j$ for exist.\ $y\in V_i$ and
            goals $g_j$ that contain $x$.
\item[C12.] (Match): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ for all $x,y$ that must match
            if $g_j$ is covered by goal $k$ in $V_i$.
\item[C13.] (If all vars in $V_i$ are distinguished, it covers only one goal):
            $v_i \land g_j \Rightarrow \neg g_k$ for appropriate views $v_i$.
\end{enumerate}
These clauses are the same clauses used by \mcdsat\ to encode QRPs \cite{arvelo:aaai06}.
In order to properly manage constant symbols, the clauses must be enhanced with:
\begin{enumerate}[C10.]
\item[C14.] (Direct inconsistency 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
\item[C15.] (Direct inconsistency 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
\item[C16.] (Direct inconsistency 3): $\neg t_{A,B}$.
\item[C17.] (Transitivity 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
\item[C18.] (Transitivity 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
\end{enumerate}
Recall that, as shown in Section~2, the main issue when handling
constants is to be sure that not two different constants are mapped
into each other either directly or indirectly.
Clauses C14--C16 prune direct inconsistent mappings, while the last
two clauses implement a restricted propagation of mappings that prune
indirect inconsistencies.
Likewise, the theory $\Delta_{com}$ that specifies complete
instantiations contains the following clauses:
\begin{enumerate}[C10.]
\item[C19.] (Cover all goals): $\bigvee_{j=1}^m g^i_j$ for $1\leq i\leq N$.
\item[C20.] (Disjunctive cover): $g^i_k \Rightarrow \neg g^j_k$ for $i\neq j$.
\item[C21.] (Prune symmetries): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
            for $1\leq j\leq m$ and $1\leq i\leq N$.
\item[C22.] (Direct inconsistency 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
\end{enumerate}
These clauses provide a sound and complete characterization of WIPs in the
sense that their models are in correspondence with the
instantiations of the workflow.
